---
title: 'DBSCAN: density-based clustering'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Given such data, k-means algorithm has difficulties for identifying theses clusters with arbitrary shape. To illustrate this situation, the following R code computes K-means algorithm on the dataset multishapes [in factoextra package]. The function fviz_cluster() [in factoextra] is used to visualize the clusters.

```{r message=FALSE, warning=FALSE}
library(factoextra)
data("multishapes")
df <- multishapes[, 1:2]
set.seed(123)
km.res <- kmeans(df, 5, nstart = 25)
fviz_cluster(km.res, df, geom = "point")
```

## R functions for DBSCAN
The function **dbscan()** [in **fpc** package] or **dbscan()** [in **dbscan** package] can be used.

```{r message=FALSE, warning=FALSE}
library(dbscan)
res.db0 <- dbscan(df, eps = 0.4, minPts = 5)
fviz_cluster(res.db0, df, geom = "point")

res.db <- dbscan(df, eps = 0.15, minPts = 5)
fviz_cluster(res.db, df, geom = "point")

table(res.db$cluster)
```
