---
title: "3.3 Data Integration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###  cbind and rbind in R
The simplest case is when we have a two datasets with either identical columns (both the number of and names) or the same number of rows. In this case, either rbind or cbind
work greatly.

#### Example 1:
```{r di1}
x1 <- c('Ali','Ahmad','Ali')
x2 <- c('Karimi','Zare','Hafezi')
x3 <- c(16,17,19)
x4 <- c(18,15,13)

D1 <- data.frame(name=x1,family=x2,math=x3)
D2 <- data.frame(name=x1,family=x2,stat=x3)

D_cbind <- cbind(D1,stat=D2$stat)
D_cbind
```

#### Example 2:
```{r di2}
D2013 <- read.csv(file="Datasets/england_payment_fulltime_2013.csv",h=T)
D2014 <- read.csv(file="Datasets/england_payment_fulltime_2014.csv",h=T)
dim(D2013)
dim(D2014)
names(D2013)
names(D2014)
head(D2013[,1:4])
head(D2014)
names(D2013)[3:4] <- paste0(names(D2013)[3:4],"_2013")
names(D2014)[4:5] <- paste0(names(D2014)[4:5],"_2014")
names(D2013)
names(D2014)
D_cbind <- cbind(D2013,D2014[,4:5])
head(D_cbind)
```

##  Joins
#### Example 3:
Consider the Example 1 and
```{r di3}
x1_1 <- c('Ahmad','Ali','Ali')
x2_1 <- c('Zare','Karimi','Hafezi')
x5 <- c(15,18,13)
D3 <- data.frame(name=x1_1,family=x2_1,stat=x5)
D3
D1
```
How we can join data frames D1 and D3?
### merge
R comes with a built-in function, called merge, to merge two data.frames.
```{r di4}
D_merge <- merge(x=D1,y=D3,by.x = c("name","family"),by.y = c("name","family"))
D_merge
```

