---
title: "The Eclat Algorithm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The Eclat algorithm is used to perform itemset mining. Itemset mining let us find frequent patterns in data like if a consumer buys milk, he also buys bread. This type of pattern is called association rules and is used in many application domains.

The basic idea for the eclat algorithm is use tidset intersections to compute the support of a candidate itemset avoiding the generation of subsets that does not exist in the prefix tree. It was originally proposed by Zaki, Parthasarathy et al.

The eclat algorithm can be found in the **arule** package of R system.

## Example 1
```{r message=FALSE, warning=FALSE}
library(arules)
## Create transaction data set.
data <- list(
  c("a","b","c"),
  c("a","b"),
  c("a","b","d"),
  c("b","e"),
  c("b","c","e"),
  c("a","d","e"),
  c("a","c"),
  c("a","b","d"),
  c("c","e"),
  c("a","b","d","e"))
data <- as(data, "transactions")
## Mine itemsets with tidLists.
f <- eclat(data, parameter = list(support = 0, tidLists = TRUE))
inspect(tidLists(f))
## Get dimensions of the tidLists.
dim(tidLists(f))
## Coerce tidLists to list.
as(tidLists(f), "list")
## Inspect visually.
image(data)
image(tidLists(f))
##Show the Frequent itemsets and respectives supports
inspect(f)
```

## Example 2
```{r message=FALSE, warning=FALSE}
library(arules)
data(Income)
inspect(Income[1:2])
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## Mine frequent itemsets with Eclat.
fsets <- eclat(Income, parameter = list(supp = 0.5))
fsets2 <- eclat(Income, parameter = list(supp = 0.1,maxlen = 15))

## Display the 5 itemsets with the highest support.
fsets.top5 <- sort(fsets[1:5])
fsets2.top5 <- sort(fsets[1:5])
inspect(fsets.top5)

## Get the itemsets as a list
as(items(fsets.top5), "list")

## Get the itemsets as a binary matrix
as(items(fsets.top5), "matrix")

## Get the itemsets as a sparse matrix, a ngCMatrix from package Matrix.
## Warning: for efficiency reasons, the ngCMatrix you get is transposed
as(items(fsets.top5), "ngCMatrix")
```

## Create rules from the itemsets
```{r echo=TRUE, message=FALSE, warning=FALSE}
rules <- ruleInduction(fsets, Income, confidence = .9)
rules <- sort(rules, by='confidence', decreasing = TRUE)
topRules <- rules[1:5]
inspect(topRules)
library(arulesViz)
plot(topRules, method="graph")
```