---
title: "Rule-Based Classification"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The C50 package
The C50 package contains an interface to the C5.0 classification model. The main two modes for this model are:<br>
<li> a basic tree-based model<br>
<li> a rule-based model

The main public resource on this model comes from the [RuleQuest website](http://www.rulequest.com/see5-info.html).

## Rule-Based Models
C5.0 can create an initial tree model then decompose the tree structure into a set of mutually exclusive rules. These rules can then be pruned and modified into a smaller set of potentially overlapping rules. The rules can be created using the rules option:

### Loading Credit dataset:
```{r message=FALSE, warning=FALSE}
credit <- read.csv("Datasets/credit.csv",h=T)
table(credit$Credit_rating)
credit$Income <- as.factor(credit$Income)
credit$Credit_cards <- as.factor(credit$Credit_cards)
credit$Education <- as.factor(credit$Education)
credit$Car_loans <- as.factor(credit$Car_loans)
set.seed(1234) 
ind <- sample(2,nrow(credit), replace=TRUE, prob=c(0.7,0.3))
credit$Credit_rating <- as.factor(credit$Credit_rating)
trainData <- credit[ind==1,]
testData <- credit[ind==2,]
```

### Building Model
```{r message=FALSE, warning=FALSE}
library(C50)  
rules <- C5.0(x = trainData[, -1],
                 y = trainData$Credit_rating,
                 rules = TRUE)
rules
summary(rules)
```

Note that no pruning was warranted for this model.
There is no plot method for rule-based models.

## Predictions
The predict method can be used to get hard class predictions or class probability estimates 
```{r message=FALSE, warning=FALSE}
yhat_rb <- predict(rules, newdata = testData)
# Confusion matrix
M <- table(testData$Credit_rating, yhat_rb)
M
```

### Sensitivity, Recall
```{r message=FALSE, warning=FALSE}
diag(prop.table(M,1))
```

### Precision
```{r message=FALSE, warning=FALSE}
diag(prop.table(M,2))
```

### Accuracy
```{r message=FALSE, warning=FALSE}
acc <- mean(testData$Credit_rating == yhat_rb)
acc
```
