---
title: "2.3 Advanced Data Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2- Create Interactive Web Graphics via 'plotly'
### Reference: <https://plot.ly/r>

### US economic time series
```{r p1, echo=TRUE, message=FALSE, warning=FALSE}
library(plotly)
data("economics")
head(economics)
attach(economics)
```

#### Description
This dataset was produced from US economic time series data available from <http://research.stlouisfed.org/fred2>. economics is in "wide" format, economics_long is in "long" format.

## Line (Time series) plots
```{r p2, echo=TRUE, message=FALSE, warning=FALSE}
plot_ly (
  x = date,
  y = unemploy,
  type = 'scatter',
  mode = 'lines' )
```

## Scatter plots
```{r p3, echo=TRUE, message=FALSE, warning=FALSE}
p <- plot_ly (
      x = pce,
      y = unemploy,
      type = 'scatter',
      mode = 'markers')
p
add_paths(p)
```

```{r p4, echo=TRUE, message=FALSE, warning=FALSE}
plot_ly(x = date, y = uempmed, color = I("red"), 
        showlegend = FALSE) %>% add_lines() %>% add_markers(color=pop)
plot_ly(x = date,y = uempmed) %>% add_ribbons(ymin = pce - 1e3, ymax = pce + 1e3)
```

Use **group_by()** to apply visual mapping once per group (e.g. one line per group)

```{r p5, echo=TRUE, message=FALSE, warning=FALSE}
data(txhousing)
head(txhousing)
txhousing %>%
group_by(txhousing$city) %>%
plot_ly(x = txhousing$date, y = txhousing$median) %>%
add_lines(color = txhousing$city)
```

```{r p6, echo=TRUE, message=FALSE, warning=FALSE}
x <- c(1, 2, 2, 1, 1, 2)
y <- c(1, 2, 2, 1, 1, 2)
z <- c(1, 1, 2, 2, 3, 3)
df <- data.frame(x = x,y = y,z = z)
plot_ly(df) %>% add_markers(x = 1.5, y = 1.5) %>% add_markers(x = x, y = y,frame = z)
```

```{r p7, echo=TRUE, message=FALSE, warning=FALSE}
p <- plot_ly(economics[1:20,]) %>% add_markers(x = ~pop,y = ~unemploy,size = ~psavert,color=~uempmed)
colorbar(p, len = 0.5)
```

## Boxplots Plots in plotly
```{r p8, echo=TRUE, message=FALSE, warning=FALSE}
plot_ly(midwest, x = ~percollege, color = ~state, type = "box")
```

## 3D Plots in plotly
```{r p9, echo=TRUE, message=FALSE, warning=FALSE}
plot_ly(z = ~volcano) %>% add_surface() 
```
