---
title: "3.2 Data Cleaning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 3.2.1 Missing Values
#### Reading the dataset with missing values
```{r dc1}
D <- read.table('Datasets/Analysis.txt',na.strings=c('XXXXXXX'))
names(D) <- c('season','size','speed','mxPH','mnO2','Cl','NO3','NH4','oPO4','PO4','Chla','a1','a2','a3','a4','a5','a6','a7')
```

#### Which objects of a dataset have at leat one missing value?
```{r dc2}
which(!complete.cases(D))
```

#### Which attributes of an object are missing?
```{r dc4}
x <- c(2,3,NA,4,NA)
is.na(x)
which(is.na(x))
sum(is.na(x))
which(is.na(D[28,]))
```

#### What percent of each attribute is missing?
```{r dc5}
n <- nrow(D)
f <- function(x) sum(is.na(x))/n
apply(D,2,f)
```

###	Ignoring the missing tuples
#### Extractig the complete objects from the dataset
```{r dc3}
D_comp <- D[complete.cases(D),]
dim(D)
dim(D_comp)
```


####Create a function that gives us the objects in dataset  have a certain number of missings.
```{r dc6}
manyNA <- function(D,a){
  n <- ncol(D)
  index <- apply(D,1,function(x) sum(is.na(x))/n > a)
  out <- list()
  out$na <- D[index,]
  out$index <- index
  out$clean <- D[!index,]
  return(out)
}
Miss_D <- manyNA(D,0.1)
Clean0.1_D <- Miss_D$clean
Miss_D$na
```

###	Use a measure of central tendency for the attribute to fill in the missing values
```{r dc7}
which(is.na(D$mxPH))
D[is.na(D$mxPH), "mxPH"] <- mean(D$mxPH, na.rm = T)
```
###	Use the most probable value to fill in the missing value
```{r dc9}
library(corrplot)
C=cor(D_comp[, 4:18])
corrplot.mixed(C, lower="number", upper="pie")
which(is.na(D$oPO4))
which(is.na(D$PO4))
D[28,'PO4']
reg <- lm(PO4 ~ oPO4, data = D_comp)
D[28,'PO4'] <- predict(reg,newdata = D[28,])
D[28,'PO4']
```

