---
title: "Loading data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Reading data from .CSV (or .TXT) files

```{r bank1}
bank <- read.csv('Datasets/customer_bank_data.csv',header = TRUE)
head(bank)
```
A flie can also be downloaded from a URL:
```{r bank3}
theUrl <- "https://raw.githubusercontent.com/haghbinh/DMwR/master/Datasets/customer_bank_data.csv"
bank1 <- read.table(file=theUrl, header=TRUE, sep=",")
```

## 2. Reading from Databases:
Databases arguably store the vast majority of the world’s data. Most of these, whether they
be PostgreSQL, MySQL, Microsoft SQL Server or Microsoft Access, can be accessed
either through various drivers, typically via an ODBC connection.
we use a simple SQLite database though these steps will be similar for most databases. 
```{r dbase}
library(RSQLite)
```
 To connect to the database we first specify the driver using dbDriver. The function’s
 main argument is the type of driver, such as “SQLite” or “ODBC.”
```{r dbase2}
drv <- dbDriver('SQLite')
```
We then establish a connection to the specific database with dbConnect. The first
argument is the driver. The most common second argument is the DSN3 connection
string for the database, or the path to the file for SQLite databases. Additional arguments
are typically the database username, password, host and port.
```{r dbase3}
con <- dbConnect(drv, 'Datasets/diamonds.db')
dbListTables(con)
dbListFields(con, name='diamonds')
dbListFields(con, name='DiamondColors')
diamondsTable <- dbGetQuery(con,
                             "SELECT * FROM diamonds",
                             stringsAsFactors=FALSE)
colorTable <- dbGetQuery(con,
                          "SELECT * FROM DiamondColors",
                          stringsAsFactors=FALSE)
longQuery <- "SELECT * FROM diamonds, DiamondColors
              WHERE 
              diamonds.color = DiamondColors.Color"
diamondsJoin <- dbGetQuery(con, longQuery,
                             stringsAsFactors=FALSE)
head(diamondsTable)
head(colorTable)
head(diamondsJoin)
```

## 3. Getting the Data from the Web
Getting the S&P 500  quotes is to use the free service
 provided by Yahoo finance, which allows you to download a CSV file with
 the quotes you want.
```{r dbase4}
library(tseries)
library(xts)
GSPC <- as.xts(get.hist.quote("^GSPC",start="1998-01-01",
                                quote=c("Open","High","Low","Close","Volume","Adjusted")))
head(GSPC)
=======
```{r bank2}
summary(bank)
```
