---
title: "Similarity and Dissimilarity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dissimilarity Measures for Nominal Attributes
“Gower's distance” is chosen by metric "gower" or automatically if some columns of x are not numeric. Also known as Gower's coefficient (1971), expressed as a dissimilarity, this implies that a particular standardisation will be applied to each variable, and the “distance” between two units is the sum of all the variable-specific distances
```{r s0}
library(cluster)
x1 <- c('a','b','a','a')
x2 <- c('y','x','x','x')
x3 <- c('e','q','e','q')
x4 <- c('z','z','z','n')
D0 <- data.frame(x1=x1,x2=x2,x3=x3,x4=x4)
d <- daisy(D0,metric = "gower")
as.matrix(d)
```

## Dissimilarity Measures for Binary Attributes
```{r s1}
library(proxy)
x <- matrix(sample(c(FALSE, TRUE), 8, rep = TRUE), ncol = 2)
x
d <- dist(x, method = "Jaccard")
as.matrix(d)
```

This is the same but less efficient
```{r s2}
as.matrix(stats::dist(x, method = "binary"))
```

### Similarity Matrix
```{r s3}
s <- simil(x, method = "Jaccard")
as.matrix(s)
```

## Dissimilarity of Numeric Data: Minkowski Distance
#### Euclidean distance:
```{r s4}
D <- iris[sample(1:150,5),-5]
d2 <- dist(D)
as.matrix(d2)
d3 <- stats::dist(D,p=3)
as.matrix(d3)
```

#### Euclidean distance(Another way):
```{r s5}
d4 <- daisy(D)
as.matrix(d4)
```

#### Manhattan distance:
```{r s6}
d5 <- daisy(D, metric = 'manhattan')
as.matrix(d5)
```

## Dissimilarity for Attributes of Mixed Types
```{r s7}
D1 <- iris[sample(1:150,5),]
D1
d6 <- daisy(D1, metric = 'gower')
as.matrix(d6)
```

## Cosine Similarity
See the following link for application Cosine Similarity in documents similarity:<br />
<http://text2vec.org/similarity.html#jaccard_similarity>
