---
title: "Principal Components Analysis for Data Reduction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Semeion Handwritten Digit
-----------------------------------------------------<br />
Data Set Characteristics: Multivariate<br />
Number of Instances: 1593<br />
Area: Computer<br />
Attribute Characteristics: Integer<br />
Number of Attributes: 256<br />
Date Donated: 2008-11-11<br />
Associated Tasks: Classification<br />
Missing Values? N/A<br />
-----------------------------------------------------
###Source
The dataset was created by Tactile Srl, Brescia, Italy 
<http://www.tattile.it/> and donated in 1994 to Semeion Research Center of Sciences of Communication, Rome, Italy <http://www.semeion.it/>, for machine learning research.

### Data Set Information:
1593 handwritten digits from around 80 persons were scanned, stretched in a rectangular box 16x16 in a gray scale of 256 values.Then each pixel of each image was scaled into a bolean (1/0) value using a fixed 
threshold.
Each person wrote on a paper all the digits from 0 to 9, twice. The commitment was to write the digit the first time in the normal way 
(trying to write each digit accurately) and the second time in a fast 
way (with no accuracy).

The best validation protocol for this dataset seems to be a 5x2CV, 50% Tune (Train +Test) and completly blind 50% Validation.

-----------------------------------------------------

### Attribute Information:
This dataset consists of 1593 records (rows) and 256 attributes 
(columns).
Each record represents a handwritten digit, orginally scanned with a 
resolution of 256 grays scale (28).
Each pixel of the each original scanned image was first stretched, and 
after scaled between 0 and 1 (setting to 0 every pixel whose value was 
under tha value 127 of the grey scale (127 included) and setting to 1 
each pixel whose orinal value in the grey scale was over 127).
Finally, each binary image was scaled again into a 16x16 square box (the 
final 256 binary attributes). 

-----------------------------------------------------

### Relevant Papers:
*M Buscema, MetaNet: The Theory of Independent Judges, in Substance Use & Misuse 33(2)1998, pp 439-461.*


```{r bank1, message=FALSE, warning=FALSE}
library(imager)
Digit <- read.table("Datasets/Digit.txt",header=TRUE)
dim(Digit)
img <- Digit[,-(257:266)]
Y <- apply(Digit[,257:266],1,function(x) sum(x*(0:9)))
i <- 81
im <- as.cimg(matrix((as.numeric(img[i,])),nr=16))
plot(im)
pca <- prcomp(img,scale =F )
dim(pca$x)
plot(pca)
pairs(pca$x[,1:4],col=Y,pch=19)
```
```{r digit2, message=FALSE, warning=FALSE}
plot(pca$x[1:100,1:2],col=Y,pch=19)
identify(pca$x[1:100,1])
im90 <- as.cimg(matrix((as.numeric(img[90,])),nr=16))
im96 <- as.cimg(matrix((as.numeric(img[96,])),nr=16))
par(mfrow=c(1,2))
plot(im90,main='image=90')
plot(im96,main='image=96')
```