---
title: "Basic Statistical Descriptions of Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Customer bank data set
```{r bank1}
bank <- read.csv('Datasets/customer_bank_data.csv',header = TRUE)
head(bank)
```
The most common and effective numeric measure of the “center” of a set of data is the (arithmetic) mean. 
```{r bank2}
mean(bank$INCOME)
```
For skewed (asymmetric) data, a better measure of the center of data is the median,
which is the middle value in a set of ordered data values.
```{r bank3}
median(bank$INCOME)
```
The 4-quantiles are the three data points that
split the data distribution into four equal parts; each part represents one-fourth of the data distribution. They are more commonly referred to as quartiles.
```{r bank4}
quantile(bank$INCOME)
```
The 100-quantiles are more commonly referred to aspercentiles; they divide the data distribution into 100 equal-sized consecutive sets.
```{r bank5}
quantile(bank$INCOME,probs = seq(0.12,0.2,by=0.01))
```
The distance between the first and third quartiles is a simple measure of spread
that gives the range covered by the middle half of the data. This distance is called the interquartile range (IQR) and is defined as<br />
IQR = Q3 − Q1
```{r bank6}
IQR(bank$INCOME)
```
## Five-Number Summary, Boxplots, and Outliers
The **five-number** summary of a distribution consists of the
median (Q2), the quartiles Q1 and Q3, and the smallest and largest individual observations, written in the order of Minimum, Q1, Median, Q3, Maximum.<br />
**Boxplots** are a popular way of visualizing a distribution.
When dealing with a moderate number of observations, it is worthwhile to plot
potential outliers individually. To do this in a boxplot, the whiskers are extended to the extreme low and high observations only if these values are less than 1.5× IQR beyond the quartiles.

```{r bank7}
boxplot(bank$AGE,col = 'cornflowerblue',main='Age')
```

Boxplots can be used in the comparisons of several sets of compatible data.

```{r bank8}
color <- c('hotpink1','greenyellow','cornflowerblue')
boxplot(bank$INCOME~bank$DEFAULTED,horizontal=T,col = color,xlab='Income',ylab='DEFAULTED')
```
## Variance and Standard Deviation
```{r bank9}
var(bank$INCOME)
sd(bank$INCOME)
```
## Shannon Entropy
The entropy quantifies the expected value of the information contained in a vector. 
```{r bank10}
library(DescTools)
Entropy(table(bank$DEFAULTED))
```
## Quantile Plot
A quantile–quantile plot, or q-q plot, graphs the quantiles of one univariate  distribution against the corresponding quantiles of another. It is a powerful visualization tool in that it allows the user to view whether there is a shift in going from one distribution to another.
```{r bank11, message=FALSE, warning=FALSE}
qqnorm(bank$INCOME, pch = 1, frame = FALSE)
qqline(bank$INCOME, col = "steelblue", lwd = 2)
```

## Histograms
 If X is nominal, the resulting graph is more commonly known as
a bar chart.
```{r bank12}
barplot(table(bank$EDUCATION),col=1:4)
barplot(table(bank$EDUCATION,bank$DEFAULTED),col=1:4,beside = TRUE)
legend(1,250,leg=unique(bank$EDUCATION),fill=1:4,border = 1:4)
```

If X is numeric, the term histogram is preferred. The subranges, referred to as buckets or bins, are disjoint subsets of the data distribution for X.

```{r bank13, message=FALSE, warning=FALSE}
hist(bank$INCOME,col = 'olivedrab2',xlab = 'Income')
hist(bank$INCOME,col = 'orange2',xlab = 'Income',probability = TRUE)
plot(density(bank$INCOME),add=TRUE,col='mediumorchid3',lwd=2)
```

## Scatter Plots 
The most effective graphical methods for determining if there
appears to be a relationship, pattern, or trend between two numeric attributes.

```{r bank14}
plot(bank$YEARSEMPLOYED, bank$OTHERDEBT,col=bank$DEFAULTED,pch=19)
legend(5,35,leg=unique(bank$DEFAULTED),fill=1:3,border = 1:3)
pairs(bank[,c(2,4,5,6)],pch=19,col='mediumorchid2')
```

## Bubble Plots 
```{r bank15}
radius <- sqrt( bank$YEARSEMPLOYED/ pi )
symbols(bank$AGE, bank$INCOME ,circles=radius,
        inches=0.35, bg="cyan", xlab="Age", ylab="Income",main='YEARSEMPLOYED')
```

## Data Correlation
```{r bank16}
cor(bank$YEARSEMPLOYED, bank$OTHERDEBT)
library(corrplot)
C=cor(iris[,-5])
corrplot.mixed(C, lower="number", upper="pie")
```
